\section{Wireless Distance Sensing and Collision Feedback}


\subsection{ESP32-C6 Network and Data Processing}
To support collision-aware robot behavior, a distributed E-Skin distance-sensing subsystem was integrated. An ESP32 microcontroller serves as the central network bridge, managing Wi-Fi and UDP communication with the wireless sensor nodes. Upon initialization via the \texttt{udp\_sensor\_task}, the ESP32 establishes a socket connection to the E-Skin interface at IP address \texttt{192.168.4.1}. It sequentially transmits initialization packets on port \texttt{17000} to establish a lock (\texttt{LOCK\_CMD\_PKT}), start the data stream (\texttt{START\_CMD\_PKT}), and configure the update transmission rate to 63~Hz (\texttt{UDR\_63HZ\_PKT}).

The continuous incoming UDP payload is received on port \texttt{17011}. Packets identified by the \texttt{0xFF} header byte are routed to a FreeRTOS queue (\texttt{eskin\_data\_queue}) for further processing. The \texttt{collision\_processing\_task} decodes the 20-byte payload to extract the sensor ID (\texttt{sc\_id}) and reconstruct the raw 16-bit proximity measurement (\texttt{prox\_raw}), scaling it to a floating-point value between 0.0 and 1.0.

While the protocol architecture supports arrays of up to 32 independent sensors, the physical proof-of-concept deployment utilized two active sensor nodes. Initial implementations exhibited state-flickering in the collision-feedback logic due to conflicting LED state requests when multiple sensors reported varying proximity levels simultaneously. To resolve this, an anti-flicker algorithm was introduced. The data processing pipeline maintains an array of the latest readings and iteratively extracts a global maximum proximity value, denoted as \texttt{max\_prox}. This scalar value is then used to define clear, unambiguous hysteresis thresholds for the system's visual feedback mechanism. Hysteresis bounds were defined at \texttt{PROXIMITY\_\allowbreak THRESHOLD\_\allowbreak CLEAR} = 0.016f (green LED, safe), \texttt{PROXIMITY\_\allowbreak THRESHOLD\_\allowbreak STOP} = 0.03f (red LED, collision warning), and \texttt{PROXIMITY\_\allowbreak THRESHOLD\_\allowbreak TOUCH} = 0.80f (blue LED, physical touch). Corresponding predefined byte arrays (e.g., \texttt{LED\_\allowbreak RED\_\allowbreak PKT}) are sent via UDP to the sensors to actuate the visual feedback without oscillation.

\subsection{Signal Handling and STM32 Integration}
For physical integration with the STM32 motion controller, the \texttt{max\_prox} value is transmitted via a UART interface. To conserve processing resources on the motion-control side, the transmission rate is throttled to 4~Hz by comparing the current FreeRTOS \texttt{xTaskGetTickCount()} against the \texttt{last\_print\_time} with a 250~ms interval. The proximity data is serialized and transmitted as a plaintext floating-point string to three decimal places, terminated by a newline character (e.g., \texttt{0.555\textbackslash n}).

The hardware integration required addressing two primary challenges. First, during initial debugging, a USB-to-UART adapter and a Python script were employed to verify data transmission. However, the ESP32 experienced voltage brownouts and failed to boot when powered directly by the adapter due to the high-current demands of the Wi-Fi initialization. This was resolved by powering the ESP32 through a separate, dedicated charging cable; consequently, only the ground (GND) and transmit (TX) lines were routed to the diagnostic UART adapter, and subsequently to the STM32, effectively isolating the power domains. Second, a UART pin conflict necessitated an alternative wiring approach. The design leverages the hardwired console TX pin (Pin D6 / GPIO16) specific to the ESP32-C6 architecture. This hardware-level decision allowed the removal of complex, resource-heavy UART driver libraries in favor of a lightweight, native \texttt{printf("\%.3f\textbackslash n", max\_prox)} function for data transmission.

On the receiving end, the STM32 integration required specific configuration within the STM32CubeMX environment. The designated UART peripheral was configured with Global Interrupts enabled to ensure reliable, non-blocking reception of the incoming data stream. The firmware buffers the incoming bytes until the \texttt{\textbackslash n} delimiter is detected. The received character array is subsequently parsed into a standard floating-point format using the \texttt{strtof} function. If the evaluated proximity exceeds the configured motion threshold, the application requests a software pause and commands the active joints to hold their measured positions. This response is not a certified emergency stop.
